---
title: "Case3 Part 1: Missing Data"
author: "Sonia Xu"
date: "November 6, 2017"
output: html_document
---

<!-- The data come from a study of malaria prevalence in Gambia. Each observation corresponds to one child who has been tested for the presence of malaria. The response Y is an indicator of whether malaria parasites were found in a blood sample from the child. The predictors are

AGE: age of the child, in years
BEDNET: indicator of whether the child has a net over his or her bed (1 = yes)
GREEN: a measure of how much greenery is around the child's village, derived from satellite images (units are arbitrary)
PHC: indicator for the presence of a public health clinic in the child's village (1 = yes)-->

#Preliminary Analysis

```{r include = F}
library(dplyr)
library(knitr)
library(ggplot2)
#read in the data set
gambia <-read.csv("gambiaMissing.csv")
```

#Missing Data: MAR
Initial exploratory data analysis reveals that the feature bednet contains missing data for 317 out of the 805. To determine whether the data is Missing at Random or Missing Completely at Random, features from the missing dataset were compared in porportion to the features from the original dataset (i.e. $ \frac{GREEN_{missing}}{GREEN_{complete}}$).     

```{r echo = F}
attach(gambia)
i <- which(is.na(BEDNET))
m_gambia <- gambia[i,]
ggplot(gambia, aes(x = unique(AGE), y = as.vector(table(AGE)))) + geom_line()


detach(gambia)
```
The missingness was skewed for any of its features