---
title: "case3_2"
author: "Sonia Xu, Vinai Oddiraju"
date: "November 13, 2017"
output: html_document
---
<!-- Week 3: Report #2 on the case study is due. This report will go significantly beyond report #1 to include initial results for different methods that were considered to analyze the data, as well as a detailed description for how the analysis was conducted and problems that arose.  The report should be well written, clear and concise while not omitting relevant details.  The roles of the authors should be described.  Class time will again be devoted to lecturing on statistical approaches relevant to the case study.--->

```{r include = F, warning = F}
knitr::opts_chunk$set(fig.width=5, fig.height=3) 
library(mice)
library(dplyr)
library(knitr)
library(ggplot2)
library(cluster)
#read in the data set
gambia <-read.csv("gambiaMissing.csv")
```
#Problem Overview:
Data from a study of malaria prevalence in Gambia was collected to assess which characteristics were significant in predicting the presence of malaria parasites in a blood sample of a child. However, missing data within the feature, $BEDNET$, an indicator of whether the child has a net over his or her bed, has created a challenge for fully analyzing and understanding the dataset. The purpose of this paper is to determine the best way to imputate the missing data in $BEDNET$ in order to create the best predictive model. Once the best fit model has been determined, the model will be used to determine to what extent each feature affects the  probability of a child containing any presence of malaria parasites in their blood sample. 
<!-- The data come from a study of malaria prevalence in Gambia. Each observation corresponds to one child who has been tested for the presence of malaria. The response Y is an indicator of whether malaria parasites were found in a blood sample from the child. The predictors are

AGE: age of the child, in years
BEDNET: indicator of whether the child has a net over his or her bed (1 = yes)
GREEN: a measure of how much greenery is around the child's village, derived from satellite images (units are arbitrary)
PHC: indicator for the presence of a public health clinic in the child's village (1 = yes)-->

#Model Overview
Before creating a robust model that predicts the presence of malaria in child's blood stream, it is important to imputate the missing data well. Three different methods in missing were tested (Clustering & 2 methods within the MICE package) and compared to assess the best way to imputate the missing data.  

#Clustering Data and Imputating Missing Data Based on Similar Characteristics

To fill in the missing data in $BEDNET$, it is assumed that children who share similar features have a simlar $BEDNET$ history. Each observation was thus clustered on their other features (i.e. $GREEN$), and separated into groups via k-means clustering with 10 groups. 10 groups were chosen because this was the highest number of groups that maintained at least 30 observations for each group, which would ensure a large enough sample size for estimation. Missing $BEDNET$ observations were imputated with the mean of each cluster.    

```{r echo = F}
code4kids <- NULL
for( i in 1:100) {
  gambia <-read.csv("gambiaMissing.csv")
  g_cluster <- kmeans(gambia[,-c(1,3)], 10, nstart = 20)
  gambia$clust <- g_cluster$cluster #add a column with clusters
  gambia <- gambia %>% group_by(clust) %>% mutate(impBEDNET = as.numeric((mean(BEDNET, na.rm = T) >= 0.5))) %>% group_by(clust) %>% mutate(impBEDNET_combine = ifelse(is.na(BEDNET), as.numeric((mean(BEDNET, na.rm = T) >= 0.5)), BEDNET))
  ind = which(!is.na(gambia$BEDNET))
  
    #Split into training and test
    training = sample_frac(gambia, .7) #Split into a training
    training_samp = as.numeric(row.names(training))
    test = gambia[-training_samp, ] #Get the test set
  
  TrMC = glm(data = training, Y ~ AGE + GREEN + PHC + impBEDNET_combine, family = "binomial")
  pred.val.c = round(predict(TrMC, test, type = "response"))
  code4kids <- rbind(code4kids, mean(test$Y == pred.val.c))
}

```
The gambia data set was split into test and training data to assess the fit of the model and missing dat imputation, where 70% of the full dataset was training and 30% was the test data. After iterating through the k-means clustering method to impute the missing data and then predicting the Y response via logit regression 100 times, the model predicted the correct test data on average, `r mean(code4kids)*100`% of the time. 

#Assessing Missing Data Imputation with Clustering
To assess the fit of the missing data clusters, the true values of the non-missing $BEDNET$ observations were compared to the predicted values of $BEDNET$ obtained from the clusters. The clusters labeled the correct $BEDNET$ value `r round(mean(gambia$BEDNET[ind] == gambia$impBEDNET[ind]) *100,2)`% of the time.   

#Using Mice Package to Imputate Data
For the methods implemented through the "mice" package, 100 iterations of 5 imputations were run in order for the estimates to converge. The estimated malaria indicator values were then compared to the real malaria indicator values to test for accuracy. Below is an overview of each method:

##Part I: Mice with Predictive Mean Matching ("PMM")
```{r echo = F}
sum(is.na(gambia$BEDNET))/nrow(gambia) #More diagnostics; just checking missing
#data 
for (col in 1:ncol(gambia)){
  print(sum(is.na(gambia[, col]))/nrow(gambia))
}
md.pattern(gambia)
FirstRun = mice(data = gambia, m = 5, meth = "pmm", maxit = 5, printFlag = F) #First imputation
Bednet_Imputations = FirstRun$imp$BEDNET #Get the imputed columns


```


```{r echo = F}
#Impute data, fit model, generate parameter estimates 
TEST_ACCURACY <- NULL
for(i in 1:100) {
  FirstRun = mice(data = gambia, m = 5, meth = "pmm", maxit = 5, printFlag = F) #First imputation
  Bednet_Imputations = FirstRun$imp$BEDNET
  #Append imputed columns to the gambia data set
  gambia$Imp1 = gambia$BEDNET
  gambia$Imp1[-ind] = Bednet_Imputations[,1]
  
  gambia$Imp2 = gambia$BEDNET
  gambia$Imp2[-ind] = Bednet_Imputations[,2]
  
  gambia$Imp3 = gambia$BEDNET
  gambia$Imp3[-ind] = Bednet_Imputations[,3]
  
  gambia$Imp4 = gambia$BEDNET
  gambia$Imp4[-ind] = Bednet_Imputations[,4]
  
  gambia$Imp5 = gambia$BEDNET
  gambia$Imp5[-ind] = Bednet_Imputations[,5]
  
  
  #Split into training and test
  training = sample_frac(gambia, .7) #Split into a training
  training_samp = as.numeric(row.names(training))
  test = gambia[-training_samp, ] #Get the test set
  #Construct 5 models on the training data
  TrM1 = glm(data = training, Y ~ AGE + GREEN + PHC + Imp1, family = "binomial")
  TrM2 = glm(data = training, Y ~ AGE + GREEN + PHC + Imp2, family = "binomial")
  TrM3 = glm(data = training, Y ~ AGE + GREEN + PHC + Imp3, family = "binomial")
  TrM4 = glm(data = training, Y ~ AGE + GREEN + PHC + Imp4, family = "binomial")
  TrM5 = glm(data = training, Y ~ AGE + GREEN + PHC + Imp5, family = "binomial")
  
  #Need to predict
  pred.val.1 = round(predict(TrM1, test, type = "response"))
  pred.val.2 = round(predict(TrM2, test, type = "response"))
  pred.val.3 = round(predict(TrM3, test, type = "response"))
  pred.val.4 = round(predict(TrM4, test, type = "response"))
  pred.val.5 = round(predict(TrM5, test, type = "response"))
  
  y_true <- test$Y
TEST_ACCURACY <- rbind(TEST_ACCURACY, c(mean(y_true == pred.val.1), mean(y_true == pred.val.2), mean(y_true == pred.val.3), mean(y_true == pred.val.4), mean(y_true == pred.val.5)))  
}

apply(TEST_ACCURACY, 2, mean)
```

#Part II: Mice with Bayesian Linear Regression
```{r echo = F}
TEST_ACCURACY2 <- NULL
for(i in 1:100) {
  FirstRun2 = mice(data = gambia, m = 5, meth = "norm", maxit = 5, printFlag = F) #First imputation
  Bednet_Imputations2 = FirstRun2$imp$BEDNET
  #Append imputed columns to the gambia data set
  gambia$Imp11 = gambia$BEDNET
  gambia$Imp11[-ind] = Bednet_Imputations2[,1]
  
  gambia$Imp21 = gambia$BEDNET
  gambia$Imp21[-ind] = Bednet_Imputations2[,2]
  
  gambia$Imp31 = gambia$BEDNET
  gambia$Imp31[-ind] = Bednet_Imputations2[,3]
  
  gambia$Imp41 = gambia$BEDNET
  gambia$Imp41[-ind] = Bednet_Imputations2[,4]
  
  gambia$Imp51 = gambia$BEDNET
  gambia$Imp51[-ind] = Bednet_Imputations2[,5]
  
  
  #Split into training and test
  training2 = sample_frac(gambia, .7) #Split into a training
  training_samp2 = as.numeric(row.names(training))
  test2 = gambia[-training_samp2, ] #Get the test set
  #Construct 5 models on the training data
  TrM12 = glm(data = training2, Y ~ AGE + GREEN + PHC + Imp11, family = "binomial")
  TrM22 = glm(data = training2, Y ~ AGE + GREEN + PHC + Imp21, family = "binomial")
  TrM32 = glm(data = training2, Y ~ AGE + GREEN + PHC + Imp31, family = "binomial")
  TrM42 = glm(data = training2, Y ~ AGE + GREEN + PHC + Imp41, family = "binomial")
  TrM52 = glm(data = training2, Y ~ AGE + GREEN + PHC + Imp51, family = "binomial")
  
  #Need to predict
  pred.val.11 = round(predict(TrM12, test, type = "response"))
  pred.val.21 = round(predict(TrM22, test, type = "response"))
  pred.val.31 = round(predict(TrM32, test, type = "response"))
  pred.val.41 = round(predict(TrM42, test, type = "response"))
  pred.val.51 = round(predict(TrM52, test, type = "response"))
  
  y_true <- test$Y
TEST_ACCURACY2 <- rbind(TEST_ACCURACY2, c(mean(y_true == pred.val.11), mean(y_true == pred.val.21), mean(y_true == pred.val.31), mean(y_true == pred.val.41), mean(y_true == pred.val.51)))  
}

apply(TEST_ACCURACY2, 2, mean)
```




