---
title: "Vinai Data Commentary"
author: "Vinai Oddiraju"
date: "November 6, 2017"
output: pdf_document
---

The data set is relatively simple. The response variable we intend to predict is labelled $Y$ and records whether or not malaria was present in a patient's blood or not, with 1 being yes and 0 being no. $Age$ denotes the age of subject. The ages were distributed across the values 1, 2, 3, and 4. Given the discrete nature of these levels, we plan to treat $Age$ as a categorical variable. The subjects were distributed roughly uniformly in age, with the most being 1. The table below displays the distribution. 

```{r, echo = FALSE, warning = FALSE}
gambia = read.csv("gambiaMissing.csv")
library(dplyr)
library(knitr)
table(gambia$AGE) %>% as.data.frame() %>% kable()
```

$PHC$ and $BEDNET$ are binary variables; $PNC$ records whether or not a public health center is nearby, and $BEDNET$ records whether or not a mosquito net was present on the child's bed at home. 

$GREEN$ is a variable that shows the percentage of greenery around the child's home. In our sample, it has 5 distinct levels, with the 40.85 being the most frequently occurring level. The table below shows the distribution.

```{r}
table(gambia$GREEN) %>% as.data.frame() %>% kable()
```

#Correlation Exploration

We wanted to test the correlations in our variables to see if any were highly correlated. Some of the correlations have been calculated below, but none of the covariates were found to be highly correlated to each other. 

```{r}
print("Age and greenery")
cor(gambia$AGE, gambia$GREEN)
print("Age and bed net")
cor(gambia$AGE, gambia$BEDNET, use = "complete.obs") 
print("Grennery and bed net")
cor(gambia$GREEN, gambia$BEDNET, use = "complete.obs")
print("PHC and bed net")
cor(gambia$PHC, gambia$BEDNET, use = "complete.obs")
print("Greenery and PHC")
cor(gambia$GREEN, gambia$PHC)
print("Age and PHC")
cor(gambia$AGE, gambia$PHC)

```

